---
title: "HireArt Case Study"
author: "Chuyi Guo"
date: "3/14/2019"
output: pdf_document
---

Data Pre-processing.

```{r, message=FALSE, warning=FALSE}
# read in the data into system.
library(readxl)
library(dplyr)
library(ggplot2)
data = read_excel('/Users/ChuyiGuo/Desktop/HireArt/HireArt - Data Analyst Exercise 10.12.17.xlsx')
data = as.data.frame(data)
# extract month, year for each line of data
data$month = as.numeric(format(as.Date(data$'Date of Contact'), "%m"))
data$year = as.numeric(format(as.Date(data$'Date of Contact'), "%Y"))
# extract some of the columns  
data = data[,c(colnames(data) %in% c('Client Name', "month", "year"))]
```

According to the spreadsheet, a client may be contacted multiply times within one month. For this situation, the client should be counted as one, based on a monthly basis. Those duplicated records need to be taken care.

```{r}
data2 = data[!duplicated(data),]
```

Get a client list for the team. That is the clients that had been contacted within those 4 years. Below shows some of the clients' name.

```{r}
client_list = unique(data2$`Client Name`)
head(client_list)
```

There are totally 'r length(client_list)' clients for this team.

```{r}
length(client_list)
```

The numbers of clients for each month of each year are calculated (named n). I.e. the team contacted 9 clients in January 2014. 

Also, the percentage of its clients for that month (named prop) are calculated. I.e. the team contacted 25% of its clients in January 2014. 

```{r}
total = as.data.frame(
  data2 %>%
    group_by(month, year) %>%
    summarise(n = n()) %>%
    mutate(prop = n/length(client_list))
)
total
```

Plot out the number of clients per month to get a intuitive overviews for this data set. It shows most of the clients are contacted in October.

```{r, fig.height=4, fig.width=8}
ggplot(total,aes(x = month, y = n )) +
  geom_point(aes(color = factor(year))) +
  geom_line(aes(color = factor(year))) +
  labs(y="Number of Clients") +
  scale_x_continuous(breaks = seq(1,12,1)) +
  scale_color_brewer(palette = "Set1")
```

Below shows the average number of clients and its percentage for each month during the four-year period. October has the highest value, which indicates on average, 80% of its clients had been contacted in October.

```{r}
month_avg = as.data.frame(
  total %>%
    group_by(month) %>%
    summarise(month_avg = mean(n), month_prop_avg = mean(prop)) 
)

ggplot(month_avg, aes(x = month, y = month_avg)) +
  geom_bar(stat="identity", position=position_dodge(),fill='skyblue') +
  geom_text(aes(label=month_avg), vjust=1.6, 
            color="black",position = position_dodge(0.9), size=3.5) +
  scale_y_continuous(sec.axis = sec_axis(~./30, name="Percentage")) +
  geom_line(aes(x = month, y = month_prop_avg*30, group=1),col='red') +
  geom_point(aes(x = month, y = month_prop_avg*30, group=1),col='red') +
  geom_text(aes(x=month, y=month_prop_avg*30,
                label=paste(round(month_prop_avg*100,digits=2),"%",sep="")),
            vjust=1.6, size=3,col='red') +
  scale_colour_brewer(palette="Set1") 
```


Therefore, it can be concluded that October is the month that the team is likely to contact the most clients.

